Identifying the L1 of non-native writers: the CMU-Haifa system\\[1em] {\large Chris Dyer$^\ast$ Manaal Faruqui$^\ast$ Noam Ordan$^\dagger$ Nathan Schneider$^\ast$\\ Yulia Tsvetkov$^\ast$ Victor Chahuneau$^\ast$ Naama Twitto$^\dagger$ Shuly Wintner$^\dagger$}\\[-3em]

### 2013-04-09 paper/paper.tex
per-feature accuracies

\section{Data}
\label{sec:data}
Our dataset in this work consists of TOEFL essays
written by speakers of eleven different L1s
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR},
distributed as part of the Native Language Identification Shared
Task \citep{tetreault-blanchard-cahill:2013:BEA}. The training data consists of 1000 essays from each native language. The essays are short, consisting of 10~to 20~sentences each. We used the provided splits of 900 documents for training and 100 for development. Each document is annotated with the author's English proficiency level (low, medium, high) and an identification (1 to 8) of the essay prompt. All essays are tokenized and split into sentences. In \Tref{tbl:stats:level} we provide some statistics on the training corpora, listed by the authors' proficiency level. All essays were tagged with the Stanford part-of-speech tagger
\citep{toutanova-03}. We did not parse the dataset.

@@

\subsection{Main Features}
\label{sec:mainfeats}
FWe use the following four feature types as the baseline features in our
model.  For features that are sensitive to frequency, we use the log of the (frequency-plus-one) as
the feature's value.  For each feature type, we report the accuracy of using just that feature type in
isolation on the training set\cjd{training? development? cross validated?} in Tab.~\ref{tbl:mainfeats}.

\begin{compactdesc}
\item[\feat{POS}] Part-of-speech $n$-grams.  Features were extracted
  to count every POS 1-, 2-, 3- and 4-gram in each
  document. 55.18\%. \ignore{CHI:IN PRP VBD, ARA:NN VBZ, KOR:DT JJ} 
\item[\feat{FreqChar}] Frequent character $n$-grams.  We experimented
  with character $n$-grams: To reduce the number of parameters, we removed features
  only those character $n$-grams that are observed more
  than $5$ times in the training corpus, and $n$ ranges from~1
    to~4. Accuracy: $74.12\%$. High-weight features include: {\tt TUR:Turk}, {\tt ITA:Ital}, {\tt JPN:Japa}.\item[\feat{CharPrompt}] Conjunction of the character $n$-gram
  features defined above with the prompt ID. Accuracy: 65.09\%.
\item[\feat{Brown}] Substitutions, deletions and insertions counts of Brown cluster unigrams and bigrams in each document. Accuracy: 72.26\%. \ignore{in top-30 features Brown features occur up to 5 times in 10 out of 11 L1s (all except Arabic)} 
\end{compactdesc}
\noindent
The accuracy of the classifier on the development set using these four
feature types is reported in \Tref{tbl:mainfeats}.

@@

\subsection{Additional Features}
To the basic set of features we now add more specific,
linguistically-motivated features, each adding a small number of
parameters to the model.  As above, we indicate the accuracy of each
feature type in isolation.

\begin{compactdesc}
\item[\feat{DocLen}] Document length in tokens. 11.81\%.
\item[\feat{Punct}] Counts of each punctuation mark. 27.41\%. 
\item[\feat{Pron}] Counts of each pronoun. 22.81\%.
\item[\feat{Position}] Positional token frequency. We use the counts
  for the first two and last three words before the period in each
  sentence as features. 53.03\%. \yt{ARA: second word `,', CHI: second word `is', HIN: second word `can'}  
\item[\feat{PsvRatio}] The proportion of passive verbs out of all
  verbs. 12.26\%.
\item[\feat{CxtFxn}] Contextual function words. Bi-grams yield
  62.79\%, tri-gram 62.32\%. \yt{CHI: some JJ, HIN: as VBN, } 
\item[\feat{Misspell}] Spelling correction edits. \sw{???}. 37.29\%. \yt{ARA: DEL:e, INS:e, SUBST:e:i; GER: SUBST:z:y; JPN: SUBST:l:r, SUBST:r:l;SPA: DOUBLE:s, MID:INS:s, INS:s }
\item[\feat{Restore}] Counts of substitutions, deletions and
  insertions of predefined tokens that we restored in the texts. 47.67\% \yt{CHI: Delete do, GER: Delete on, ITA Delete be}  
\end{compactdesc}
\noindent
\Tref{tbl:addfeats} reports the empirical improvement that each of
these brings independently when added to the main features
(\Sref{sec:mainfeats}).

\begin{table}[hbt]
\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\textsc{main} + \feat{Position} & 6,153,015 & 81.00 & 1.0 \\
\textsc{main} + \feat{PsvRatio} & 5,664,472 & 81.00 & 1.0 \\
\textsc{main} & 5,664,461 & 81.09 & 1.0 \\
\textsc{main} + \feat{DocLen} & 5,664,472 & 81.09 & 1.0 \\
\textsc{main} + \feat{Pron} & 5,664,736 & 81.09 & 1.0 \\
\textsc{main} + \feat{Punct} & 5,664,604 & 81.09 & 1.0 \\
\textsc{main} + \feat{Misspell} & 5,799,860 & 81.27 & 5.0 \\
\textsc{main} + \feat{Restore} & 5,682,589 & 81.36 & 5.0 \\
\textsc{main} + \feat{CxtFxn} & 7,669,684 & 81.73 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} features plus additional feature groups, added independently. 
$\ell_2$ regularization was tuned as in \Tref{tbl:mainfeats} (two values, 1.0 and 5.0, were tried for each 
configuration; more careful tuning might produce slightly better accuracy).
Results are sorted by accuracy; only three groups exhibited independent improvements over the \textsc{main} feature set.}
\label{tbl:addfeats}
\end{table}

\subsection{Discarded Features}
We also tried several other feature types that did not improve the
accuracy of the classifier on the development set.
\begin{compactdesc}
\item[Cohesive markers] Counts of each cohesive marker. 25.71\%.
\item[Cohesive verbs] Counts of each cohesive verb. 22.85\%.
\item[Function words] Counts of function words. 42.47\%. This feature
  is subsumed by the highly discriminative \feat{CxtFxn} feature.
\end{compactdesc}

\section{Results}
\label{sec:results}
The full model that we used to classify the test set combines all
features listed in \Tref{tbl:addfeats}. Using all these features, the
accuracy on the development set is~84.55\%, and on the test set it
is~81.5\%. \Tref{tbl:matrix} lists the confusion matrix on the test
set, as well as precision, recall and $F_1$-score for each L1.
The largest class of errors was predicting Telugu where the correct label 
was Hindi\nss{or vice versa?}---this happened 18 times.

>>>>

\section{Data}
\label{sec:data}
Our dataset in this work consists of TOEFL essays
written by speakers of eleven different L1s
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR},
distributed as part of the Native Language Identification Shared
Task \citep{tetreault-blanchard-cahill:2013:BEA}. The training data consists of 1000 essays from each native language. The essays are short, consisting of 10~to 20~sentences each. We used the provided splits of 900 documents for training and 100 for development. Each document is annotated with the author's English proficiency level (low, medium, high) and an identification (1 to 8) of the essay prompt. All essays are tokenized and split into sentences. In \tref{tbl:stats:level} we provide some statistics on the training corpora, listed by the authors' proficiency level. All essays were tagged with the Stanford part-of-speech tagger
\citep{toutanova-03}. We did not parse the dataset.

@@

\begin{table}\centering\small
\begin{tabular}{l@{}c}
\bf Feature & \bf Accuracy (\%) \\
\hline
\feat{POS} & 55.18\\
\feat{FreqChar} & 74.12\\
\feat{CharPrompt} & 65.09\\
\feat{Brown} & 72.26\\
\hline
\feat{DocLen} & 11.81\\
\feat{Punct} & 27.41\\
\feat{Pron} & 22.81\\
\feat{Position} & 53.03\\
\feat{PsvRatio} & 12.26\\
\feat{CxtFxn} \textsmaller{(bigram)} & 62.79\\
\feat{CxtFxn} \textsmaller{(trigram)} & 62.32\\
\feat{Misspell} & 37.29\\
\feat{Restore} & 47.67\\ 
\hline
\feat{CohMark} & 25.71\\
\feat{CohVerb} & 22.85\\
\feat{FxnWord} & 42.47\\
\hline
\end{tabular}
\caption{Independent performance of feature types detailed in \Sref{sec:mainfeats}, \Sref{sec:addfeats} and \Sref{sec:discarded}. 
Accuracy is averaged over 10 folds of cross-validation on the training set.}
\end{table}

\subsection{Main Features}
\label{sec:mainfeats}
FWe use the following four feature types as the baseline features in our
model.  For features that are sensitive to frequency, we use the log of the (frequency-plus-one) as
the feature's value.  \Tref{tbl:mainfeats} reports the accuracy of using each feature type in
isolation (with 10-fold cross-validation on the training set).

\begin{compactdesc}
\item[\feat{POS}] Part-of-speech $n$-grams.  Features were extracted
  to count every POS 1-, 2-, 3- and 4-gram in each
  document. \ignore{CHI:IN PRP VBD, ARA:NN VBZ, KOR:DT JJ} 
\item[\feat{FreqChar}] Frequent character $n$-grams.  We experimented
  with character $n$-grams: To reduce the number of parameters, we removed features
  only those character $n$-grams that are observed more
  than $5$ times in the training corpus, and $n$ ranges from~1
    to~4. High-weight features include: {\tt TUR:Turk}, {\tt ITA:Ital}, {\tt JPN:Japa}.
\item[\feat{CharPrompt}] Conjunction of the character $n$-gram
  features defined above with the prompt ID.
\item[\feat{Brown}] Substitutions, deletions and insertions counts of Brown cluster 
unigrams and bigrams in each document.\ignore{in top-30 features Brown features occur up to 5 times in 10 out of 11 L1s (all except Arabic)} 
\end{compactdesc}
\noindent
The accuracy of the classifier on the development set using these four
feature types is reported in \tref{tbl:mainfeats}.

@@

\subsection{Additional Features}\label{sec:addfeats}
To the basic set of features we now add more specific,
linguistically-motivated features, each adding a small number of
parameters to the model.  As above, we indicate the accuracy of each
feature type in isolation.

\begin{compactdesc}
\item[\feat{DocLen}] Document length in tokens.
\item[\feat{Punct}] Counts of each punctuation mark. 
\item[\feat{Pron}] Counts of each pronoun.
\item[\feat{Position}] Positional token frequency. We use the counts
  for the first two and last three words before the period in each
  sentence as features. \yt{ARA: second word `,', CHI: second word `is', HIN: second word `can'}  
\item[\feat{PsvRatio}] The proportion of passive verbs out of all
  verbs.
\item[\feat{CxtFxn}] Contextual function words. \yt{CHI: some JJ, HIN: as VBN, } 
\item[\feat{Misspell}] Spelling correction edits. \sw{???}. \yt{ARA: DEL:e, INS:e, SUBST:e:i; GER: SUBST:z:y; JPN: SUBST:l:r, SUBST:r:l;SPA: DOUBLE:s, MID:INS:s, INS:s }
\item[\feat{Restore}] Counts of substitutions, deletions and
  insertions of predefined tokens that we restored in the texts. \yt{CHI: Delete do, GER: Delete on, ITA Delete be}  
\end{compactdesc}
\noindent
\Tref{tbl:addfeats} reports the empirical improvement that each of
these brings independently when added to the main features
(\Sref{sec:mainfeats}).

\begin{table}[hbt]
\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\textsc{main} + \feat{Position} & 6,153,015 & 81.00 & 1.0 \\
\textsc{main} + \feat{PsvRatio} & 5,664,472 & 81.00 & 1.0 \\
\textsc{main} & 5,664,461 & 81.09 & 1.0 \\
\textsc{main} + \feat{DocLen} & 5,664,472 & 81.09 & 1.0 \\
\textsc{main} + \feat{Pron} & 5,664,736 & 81.09 & 1.0 \\
\textsc{main} + \feat{Punct} & 5,664,604 & 81.09 & 1.0 \\
\textsc{main} + \feat{Misspell} & 5,799,860 & 81.27 & 5.0 \\
\textsc{main} + \feat{Restore} & 5,682,589 & 81.36 & 5.0 \\
\textsc{main} + \feat{CxtFxn} & 7,669,684 & 81.73 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} features plus additional feature groups, added independently. 
$\ell_2$ regularization was tuned as in \tref{tbl:mainfeats} (two values, 1.0 and 5.0, were tried for each 
configuration; more careful tuning might produce slightly better accuracy).
Results are sorted by accuracy; only three groups exhibited independent improvements over the \textsc{main} feature set.}
\label{tbl:addfeats}
\end{table}

\subsection{Discarded Features}\label{sec:discarded}
We also tried several other feature types that did not improve the
accuracy of the classifier on the development set.
\begin{compactdesc}
\item[\feat{CohMark}] Counts of each cohesive marker.
\item[\feat{CohVerb}] Counts of each cohesive verb.
\item[\feat{FxnWord}] Counts of function words. These features
  are subsumed by the highly discriminative \feat{CxtFxn} features.
\end{compactdesc}

\section{Results}
\label{sec:results}
The full model that we used to classify the test set combines all
features listed in \tref{tbl:addfeats}. Using all these features, the
accuracy on the development set is~84.55\%, and on the test set it
is~81.5\%. \Tref{tbl:matrix} lists the confusion matrix on the test
set, as well as precision, recall and $F_1$-score for each L1.
The largest class of errors was predicting Telugu where the correct label 
was Hindi\nss{or vice versa?}---this happened 18 times.
